Dataset properties

### 2025-03-25 paper.tex
Fix merging

Suppose that $K$ classes $C_1, \ldots, C_K$ are distributed according to probability distributions $p_i$ on a  space $X$. Let us denote the $r_{ij}(x)$ the output of Bayes (binary) classifier for $x$ in $X$ yielding the probability $p(x \in C_i \mid x\in C_i \cup C_j )$. Then we have
$$
r_{ij}(x)= \frac {p_i(x)}{p_i(x) + p_j(x)}.
$$

The system of equations is called Bradley--Terry equations.  The equations in \eqref{eq:bt}  can be transformed to a system of linear equations. The theoretical basis for one-vs-one classification framework is provided by the following result.

\begin{prop}
Suppose $p_i(x) >0$ so that all $r_{ij}(x)$ are defined. The system of linear equations \eqref{eq:bt}  has a unique solution which is the multi-class Bayes classifier.
\end{prop}
\begin{proof}
TODO	
\end{proof}

Suppose we have estimators $\hat r_{ij}$ for Bayes predictions $r_{ij}$ for all pairs $i\not= j$. A \emph{coupling method} obtains a multi-class probabilistic estimate $\boldsymbol{\hat p}= (\hat p_1, \ldots, \hat p_K)$ that satisfy
\begin{equation}
	\hat{r}_{ij}(x) \approx \frac {\hat p_i(x)}{\hat p_i(x) + \hat p_j(x)}. \label{eq:bt}
\end{equation}

@@

\subsection{Desiderata on coupling method}

@@

\emph{Bradley-Terry consistency} refers to the outcome of a procedure providing a unique solution. The requirement states that shouldwe apply the coupling procedure to binary Bayes classifiers, then the result is the output of multi-class Bayes classifier i.e. satisfies \eqref{eq:bt}.

@@

@@

>>>>

Suppose that $K$ classes $C_1, \ldots, C_K$ are distributed according to probability distributions $p_i$ on a  space $X$. Let us denote the $r_{ij}(x)$ the output of Bayes (binary) classifier for $x$ in $X$ yielding the probability $p(x \in C_i \mid x\in C_i \cup C_j )$. Then we have
\begin{equation}
r_{ij}(x)= \frac {p_i(x)}{p_i(x) + p_j(x)}. \label{eq:bt1}
\end{equation}

The system of equations is called Bradley--Terry equations.  The equations in \eqref{eq:bt1}  can be transformed to a system of linear equations. The theoretical basis for one-vs-one classification framework is provided by the following result.

\begin{prop}
Suppose $p_i(x) >0$ so that all $r_{ij}(x)$ are defined. The system of linear equations \eqref{eq:bt1}  has a unique solution which is the multi-class Bayes classifier.
\end{prop}
\begin{proof}
TODO	
\end{proof}

Suppose we have estimators $\hat r_{ij}$ for Bayes predictions $r_{ij}$ for all pairs $i\not= j$. A \emph{coupling method} obtains a multi-class probabilistic estimate $\boldsymbol{\hat p}= (\hat p_1, \ldots, \hat p_K)$ that satisfy
\begin{equation}
	\hat{r}_{ij}(x) \approx \frac {\hat p_i(x)}{\hat p_i(x) + \hat p_j(x)}. \label{eq:bt2}
\end{equation}

@@

\subsection{Exact desiderata on coupling method}

@@

\emph{Bradley-Terry consistency} refers to the outcome of a procedure providing a unique solution. The requirement states that should we apply the coupling procedure to binary Bayes classifiers, then the result is the output of multi-class Bayes classifier i.e. satisfies \eqref{eq:bt1}.

@@

\subsection{Inexact desiderata for coupling methods}

@@

\section{Examples of coupling methods}

\subsection{Arboreal coupling methods}

This coupling method ignores all but $K-1$ of the pairwise predictions. A simple example for 3 classes is the classifier which ignores prediction of the binary classifier distinguishing between classes 1 and 3. The two equations form \eqref{eq:bt2} together with the requirement that the prediction probabilities should sum to 1 yield the following system of equations
\begin{equation}
\begin{split}
\frac{\hat p_1}{\hat p_2} &= \vec{r}_{12}\\
\frac{\hat p_2}{\hat p_3} &= \vec{r}_{23}\\
\hat p_1 + \hat p_2 + \hat p_3 &= 0
\end{split}
\end{equation}

For $K=3$, there are three different arboreal coupling methods, which we shall call $omit12$, $omit13$, $omit23$.

In order to obtain a regular system of equations for  larger values of $K$ , the resulting classifiers must form a tree within the complete graph on all classes.

This method is non-canonical, because there is no natural way to choose a tree in a complete graph.

\subsection{Hinton's oracle}

This oracle method assumes it has the access to the ground truth $y = C_i$. It outputs probabilities by solving only the equations
$$
\frac{p_i}{p_j} = \vec{r}_{ij},\quad \textrm{for }i\not = j,
$$
together  with total sum equation
$$
\sum_{i=1}^K  \hat p_i = 1.
$$

It is thus an arboreal method where the underlying tree forms a star. This method is clearly the optimal one with respect to Hinton's minimality criterion. It is not a true classification method, because it ``peeks'' at the label. Its main utility is to provide a way to benchmark other methods with respect to Hinton's minimality.

%\subsection{Radial coupling}
%
%This coupling method aims to convert Hinton's oracle to a practical method. Denote by $\hat{\boldsymbol{h}}_i$ the prediction of Hinton's oracle method, if the true class was $C_i$. These predictions will generally span the simplex $\sum_i \hat p_i = 1$. In that case the prediction $\boldsymbol{r}$ of the method  can be expressed as a linear combination of $\hat{\boldsymbol{h}}_i$. The radial coupling method requires that for  weights we could in fact use the components of $\boldsymbol{r}$ i.e.
%\begin{equation}
%	\boldsymbol{r} = \sum_{i=1}^K r_i \hat{\boldsymbol{h}}_i. \label{eq:radial}
%\end{equation}
%
%The intuition behind \eqref{eq:radial} is that if some component of $\boldsymbol{r}$ is close to 1 (and thus the remaining components are close to zero), then the prediction $\boldsymbol{r}$ should be close to the prediction of Hinton's oracle $\hat{\boldsymbol{h}}_i$.

\subsection{Normal coupling}

\subsection{Wu-Lin-Weng's method}

### 2025-03-25 quarto/data.qmd
Fix merging

df <- map(files, function(f) {
# df <- map("dna", function(f) {
    dfs <- read_wlws(800, f, 0)  
    data.frame(dataset = f, K = max(dfs$train$class_id))
}) |> list_rbind()

@@

        tot <- 0 
        for (i in 1:(K-1)) {
            for (j in (i+1):K) {
                X1 <- X[dft$class_id == i,]
                X2 <- X[dft$class_id == j,]
                objective <- Minimize(sum_squares(w))
                constraints <- list (X1 %*% w + b <= -1,
                                     X2 %*% w + b >= 1)
                problem  <- Problem(objective, 
                        constraints = constraints)
                result <- solve(problem, "ECOS")
                if(result$status == "optimal") {
                    tot <- tot + 1
                } else {
                    if(result$status != "infeasible") {
                        print(result$status)
                    }
                }
            }
        }
#        stopifnot(F)
        data.frame(dataset = df$dataset[r], K = K, tot = tot)
    }) |> list_rbind()
}) |> list_rbind()
```

>>>>

# df <- map(files[1], function(f) {
df <- map("dna", function(f) {
    dfs <- read_wlws(800, f, 0)  
    data.frame(dataset = f, K = max(dfs$train$class_id))
}) |> list_rbind()

@@

        tot <- 0 
        for (i in 1:(K-1)) {
            for (j in (i+1):K) {
                X1 <- X[dft$class_id == i,]
                X2 <- X[dft$class_id == j,]
                objective <- Minimize(sum_squares(w))
                constraints <- list (X1 %*% w + b <= -1,
                                     X2 %*% w + b >= 1)
                problem  <- Problem(objective, 
                        constraints = constraints)
                result <- solve(problem, "ECOS")
                if(result$status == "optimal") {
                    tot <- tot + 1
                } else {
                    if(result$status != "infeasible") {
                        print(result$status)
                    }
 #                 stopifnot(result$status == "infeasible")

                }
            }
        }
#        stopifnot(F)
        data.frame(dataset = df$dataset[r], K = K, tot = tot)
    }) |> list_rbind()
}) |> list_rbind()
```
